Analysing the effect of land-use type on the proportion of scouts

### 2022-02-08 analysis/fit_models_durations.Rmd
reran with export to SVG
updated main figures with loading boxplots

REMEBER: Once we know how we are doing the conversion from duration to distance, or what we are doing with it, reset this branch to the remote one and check everthing is as was. Then make the edits in the dev branch, merge the relevent bits in properly and then redo the analysis, remaking the figures for the code. You MUST reset this to remote to prevent overwiting graphs and what not.

@@

In this document I have fit the collective and individual model to each site. So al the sites have reasonable likelihood values, however in some cases the fits are simply wrong (i.e. they show the probability of sampling a value greater than on equal to $x$ as increasing in the tail). I need to trouble shoot these with Vincent. I also need to get some advice on how to plot the model fits to the histogram of data.

Would be a good idea to knit this as a single document / run in R studio on the laptop.

```{r}

calc_dist_actual <- function(duration) {
  m = 1.38
  b = 0. #0.17
  distance = (duration-b) / m
  # distance = (1/m)*duration-(b/m)
  return(distance)
}

data <- read.csv("data/FullHBForagingData.csv")

head(data)

data$newforaging_distance <- calc_dist_actual(data$duration.seconds)

head(data)

par(mfrow = c(2, 2))
hist(data$duration.seconds)
hist(data$ForagingDistance_km)
hist(data$newforaging_distance)

min(data$newforaging_distance)
min(data$duration.seconds)

alldata <- data %>%
  select(date, site, duration.seconds) %>%
  rename(foraging_distance = duration.seconds)

head(alldata)

```

@@

@@

ggsave(
  plot = stu_zsl_fit,
  filename = "results/figures/stu_zsl_fit.svg",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 100
)

```

>>>>

@@

```{r}
# set waggle dance duration in seconds as foraging distance for analysis
data <- read.csv("data/FullHBForagingData.csv")

alldata <- data %>%
  select(date, site, duration.seconds) %>%
  rename(foraging_distance = duration.seconds)
```

@@

ggsave(
  plot = sites_model_plot,
  filename = "results/figures/sites_model_plot.svg",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = stu_zsl_fit,
  filename = "results/figures/stu_zsl_fit.svg",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 100
)
```

### 2022-02-08 analysis/landuse_effect_jackknife.Rmd
updated main figures with loading boxplots

# process results of jackknife for further analysis
urban_jk_loadings <- do.call(bind_rows, jkloadings)
urban_jk_pls <- do.call(bind_rows, jkplsdata)
urban_jk_coef <- do.call(bind_rows, jkcoefs)
```

@@

# process results of jackknife for further analysis
rural_jk_loadings <- do.call(bind_rows, jkloadings)
rural_jk_pls <- do.call(bind_rows, jkplsdata)
rural_jk_coef <- do.call(bind_rows, jkcoefs)
```

>>>>

# process results of jackknife for further analysis
urban_jk_loadings <- do.call(bind_rows, jkloadings)
urban_jk_loadings$cutoff <- sqrt(
  1 / (length(unique(urban_jk_loadings$landuse_type)) - 1)
)
urban_jk_pls <- do.call(bind_rows, jkplsdata)
urban_jk_coef <- do.call(bind_rows, jkcoefs)
```

@@

# process results of jackknife for further analysis
rural_jk_loadings <- do.call(bind_rows, jkloadings)
rural_jk_loadings$cutoff <- sqrt(
  1 / (length(unique(rural_jk_loadings$landuse_type)) - 1)
)
save(urban_jk_loadings, rural_jk_loadings, file = "results/pls-loadings.RData")
rural_jk_pls <- do.call(bind_rows, jkplsdata)
rural_jk_coef <- do.call(bind_rows, jkcoefs)
```

### 2022-02-08 analysis/landuse_effect_prop_scouts.Rmd
reran with export to SVG
updated main figures with loading boxplots

print(model$pp[, 1:3])
sqrt(1 / (ncol(urban.pcadata) - 1))

# when multiple PCs have the best fit, summing them gives the overall effect of
# the PCs, just as we do in the model formula.
rowSums(model$pp[, 1:3])
```

@@

```{r, rural_pls_loadings, message = FALSE, warning = FALSE}
agrirural_loadings <- list(
  loadings = model$pp[, 1],
  cutoff = sqrt(1 / (ncol(rural.pcadata) - 1))
)
save(agrirural_loadings, urban_loadings, file = "results/pls-loadings.RData")

@@

**Stats notes: things to add or improve in the stats**

* One thing it would be nice to do, but which I have not figured out how to do, is to see how much od the data the PLS PCs explain. With the PCA we obtain the variance a PC explains in the data: i.e. the variance in land-use types within our data ($R^2_x$). A regression after then tells us how much of the variance in the proportion of scouts that particular PC explains ($R^2_y$). The PLS optimise a PC to describe as much of the variation in the proportion of scouts as possible ($R^2_y$) but it would be also nice to know how much of the variation in land-use type ($R^2_x$) that PC explains.

* It would be nice to add confidence intervals to the model fit plots. Unsure how to do this with a beta regression and not all that sure it is that important?

# Figures, results write up and main discussion points for manuscript

@@

# make model fit graph of pc1 against prop scouts and mark the sites
ar_fit <- ggplot(
  data = ar_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = rural_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  geom_point(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(textsize)
  ) +
  theme_classic() +
  theme(legend.position = "none", text = element_text(size=rel(textsize-2))) +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

@@

# make loadings plot
argi.rural.loadings <- tibble(
  "Landuse" = names(agrirural_loadings$loadings),
  "Correlation" = agrirural_loadings$loadings
) %>%
  mutate(
    significant = ifelse(
      Correlation > agrirural_loadings$cutoff,
      Landuse,
      ifelse(
        Correlation < -agrirural_loadings$cutoff,
        Landuse,
        "Other"
      )
    )
  ) %>%
  mutate(
    Landuse = ifelse(
      Landuse == "Non.agricultural.managed.green.space",
      "NAMGS",
      ifelse(
        Landuse == "Non.agricultural.unmanaged.green.space",
        "NAUMGS",
        Landuse
      )
    )
  )

loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(textsize)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(0, -0.05), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size=rel(textsize-2))
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  ylim(-0.6, 0.6)

@@

# MAPS BELLOW PLOT
# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[1]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[2]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[3]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  nrow = 1,
  align = "h",
  labels = c(
    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
  ),
  label_size = rel(textsize*5)
)

mfitpls <- plot_grid(
  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  sitelumaps,
  ncol = 1,
  align = "l",
  rel_heights = c(1, .5),
  labels = c(
    "A", ""
  ),
  label_size = rel(textsize*5)
)

# all together
comb <- plot_grid(
  mfitpls, loading_plot,
  nrow = 1, align = "h",
  rel_widths = c(1, .4),
  labels = c("", "B", ""),
  label_size = rel(textsize*5)
)

@@

ggsave(
  plot = comb,
  filename = "results/figures/lu_prop_scouts_agrirural.svg",
  width = 183,
  height = 183,
  units = "mm",
  dpi = 300
)

```

@@

# make model fit graph of pc1 against prop scouts and mark the sites with the
# lowest, middle and highest proportion of scouts
u_fit <- ggplot(
  data = u_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_point(
    data = u_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = u_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(textsize)
  ) +
  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = urban_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  theme_classic() +
  theme(legend.position = "none", text = element_text(size=rel(textsize-2))) +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

@@

# make loadings plot
urban.loadings <- tibble(
  "Landuse" = names(urban_loadings$loadings),
  "Correlation" = urban_loadings$loadings
) %>%
  mutate(
    significant = ifelse(
      Correlation > urban_loadings$cutoff,
      Landuse,
      ifelse(
        Correlation < -urban_loadings$cutoff,
        Landuse,
        "Other"
      )
    )
  )

loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(textsize)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.3), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  theme(text = element_text(size = 10)) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -urban_loadings$cutoff,
      ymax = urban_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size=rel(textsize-2))
  ) +
  scale_colour_manual(
    values = c(
      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
      "parks.allotments.cemeteries" = "#228833",
      "water" = "#BBBBBB", other = "#BBBBBB"
    )
  ) +
  scale_y_continuous(breaks = seq(-1, 1, by = 0.4), limits = c(-1, 1))

@@

# MAPS BELLOW PLOT
# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[1]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[2]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[3]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  nrow = 1,
  align = "h",
  labels = c(
    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
  ),
  label_size = rel(textsize*5)
)

mfitpls <- plot_grid(
  u_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  sitelumaps,
  ncol = 1,
  align = "l",
  rel_heights = c(1, .5),
  labels = c(
    "A", ""
  ),
  label_size = rel(textsize*5)
)

# all together
comb <- plot_grid(
  mfitpls, loading_plot,
  nrow = 1, align = "h",
  rel_widths = c(1, .45),
  labels = c("", "B", ""),
  label_size = rel(textsize*5)
)

@@

```

@@

unique(vals)

```

>>>>

print(model$pp[, 1:3])
sqrt(1 / (ncol(urban.pcadata) - 1))
```

@@

```{r, rural_pls_loadings, message = FALSE, warning = FALSE}
agrirural_loadings <- list(
  loadings = model$pp[, 1],
  cutoff = sqrt(1 / (ncol(rural.pcadata) - 1))
)

@@

# Figures, results write up and main discussion points

@@

# make model fit graph of pc1 against prop scouts and mark the sites
ar_fit <- ggplot(
  data = ar_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = rural_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  geom_point(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(textsize)
  ) +
  theme_classic() +
  theme(legend.position = "none", text = element_text(size = rel(textsize - 2))) +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

@@

# make loadings plot
argi.rural.loadings <- rural_jk_loadings %>%
  rename(
    Landuse = landuse_type,
    Correlation = loadings
  ) %>%
  mutate(
    Landuse = tools::toTitleCase(
      gsub("\\.", " ", rural_jk_loadings$Landuse)
    ),
    significant = ifelse(
      Correlation > rural_jk_loadings$cutoff,
      Landuse,
      ifelse(
        Correlation < -rural_jk_loadings$cutoff,
        Landuse,
        "Other"
      )
    )
  ) %>%
  mutate(
    Landuse = ifelse(
      Landuse == "Non.agricultural.managed.green.space",
      "NAMGS",
      ifelse(
        Landuse == "Non.agricultural.unmanaged.green.space",
        "NAUMGS",
        Landuse
      )
    )
  )

rural_loading_text <- tibble(
  landuse_type = unique(argi.rural.loadings$Loadings), result = 0
)
loading_plot <- ggplot(data = argi.rural.loadings, aes(
  x = Landuse, y = Correlation, fill = Landuse
)) +
  geom_boxplot() +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -rural_cuttoff,
    ymax = rural_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  )

@@

# MAPS BELLOW PLOT
# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[1]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[2]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[3]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  nrow = 1,
  align = "h",
  labels = c(
    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
  ),
  label_size = rel(textsize * 5)
)

mfitpls <- plot_grid(
  ar_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  sitelumaps,
  ncol = 1,
  align = "l",
  rel_heights = c(1, .5),
  labels = c(
    "A", ""
  ),
  label_size = rel(textsize * 5)
)

# all together
comb <- plot_grid(
  mfitpls, loading_plot,
  nrow = 1, align = "h",
  rel_widths = c(1, .4),
  labels = c("", "B", ""),
  label_size = rel(textsize * 5)
)

@@

ggsave(
  plot = comb,
  filename = "results/figures/lu_prop_scouts_agrirural.svg",
  width = 183,
  height = 183,
  units = "mm",
  dpi = 300
)
```

@@

# make model fit graph of pc1 against prop scouts and mark the sites with the
# lowest, middle and highest proportion of scouts
u_fit <- ggplot(
  data = u_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_point(
    data = u_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = u_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(textsize)
  ) +
  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = urban_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  theme_classic() +
  theme(legend.position = "none", text = element_text(size = rel(textsize - 2))) +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

@@

# make loadings plot
urban.loadings <- urban_jk_loadings %>%
  rename(
    Landuse = landuse_type,
    Correlation = loadings
  ) %>%
  mutate(
    Landuse = tools::toTitleCase(
      gsub("\\.", " ", urban_jk_loadings$Landuse)
    ),
    significant = ifelse(
      Correlation > urban_jk_loadings$cutoff,
      Landuse,
      ifelse(
        Correlation < -urban_jk_loadings$cutoff,
        Landuse,
        "Other"
      )
    )
  )

urban_loading_text <- tibble(
  landuse_type = unique(urban.loadings$Landuse), result = 0
)
loading_plot <- ggplot(data = urban.loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_boxplot() +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
      "Sparse Residential" = "#EE6677", "Railway" = "#CCBB44",
      "Parks Allotments Cemeteries" = "#228833",
      "Water" = "#BBBBBB", Other = "#BBBBBB"
    )
  )

@@

# MAPS BELLOW PLOT
# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[1]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[2]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  site_plots[[3]] + theme(
    legend.position = "none",
    plot.margin = unit(c(0., 0., -1.2, 0.), "cm")
  ),
  nrow = 1,
  align = "h",
  labels = c(
    names(site_plots)[1], names(site_plots)[2], names(site_plots)[3]
  ),
  label_size = rel(textsize * 5)
)

mfitpls <- plot_grid(
  u_fit + theme(plot.margin = unit(c(0.2, 0.2, 1.2, 0.2), "cm")),
  sitelumaps,
  ncol = 1,
  align = "l",
  rel_heights = c(1, .5),
  labels = c(
    "A", ""
  ),
  label_size = rel(textsize * 5)
)

# all together
comb <- plot_grid(
  mfitpls, loading_plot,
  nrow = 1, align = "h",
  rel_widths = c(1, .45),
  labels = c("", "B", ""),
  label_size = rel(textsize * 5)
)

@@

ggsave(
  plot = comb,
  filename = "results/figures/lu_prop_scouts_urban.svg",
  width = 183,
  height = 183,
  units = "mm",
  dpi = 300
)
```

@@

unique(vals)
```
